---
title: "Visualization Project"
author: "Wentao Wu"
date: '2022-09-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, fig.width=8.1, fig.height=20, out.width = "70%", fig.align = "center")
```

```{r, echo=FALSE, include = FALSE}
library(fivethirtyeight) 
library(ggplot2)
library(dplyr) 
library(tidyr)
library(forcats)
```

```{r pressure, echo=FALSE}
data("bad_drivers")
FC_data = bad_drivers %>% group_by(state) %>%
  summarise(num_drivers = num_drivers,
            num_speeding = perc_speeding/100 * num_drivers, 
            num_no_speeding = num_drivers - num_speeding,
            num_alchol = perc_alcohol/100 * num_drivers,
            num_no_alchol = num_drivers - num_alchol,
            num_no_distracted = perc_not_distracted/100 * num_drivers,
            num_distracted = num_drivers - num_no_distracted,
            num_no_previous = perc_no_previous/100 * num_drivers,
            num_previous = num_drivers - num_no_previous,
            premium = insurance_premiums,
            loss = losses
  ) %>%
  gather("type","value", 2:12)
```

# Drivers Involved In Fatal Collisions Who Were Not Distracted
```{r, echo=FALSE}
ND_data = FC_data %>% filter(type == "num_no_distracted" | type == "num_distracted")
ND_color = c(num_no_distracted = "#E3A834", num_distracted = "#E3D99D")

ggplot(data = ND_data, aes(x = fct_rev(state), y = value, fill = type)) + 
  coord_flip() + 
  labs(x= NULL, y = NULL) + 
  geom_col(width = 0.9) + 
  ggtitle("Drivers Involved In Fatal Collisions Who Had Not \nBeen Involved In Any Previous Accidents", 
          subtitle = "As a share of the number of fatal collisions per billion miles, 2012") + 
  scale_fill_manual(labels = c("NON-DISTRACTED COLLISIONS PER BILLION MILES", "TOTAL COLLISIONS PER BILLION MILES"), 
                    values = ND_color, name = NULL) + 
  theme(legend.direction = "vertical", 
        legend.position = "top",
        legend.justification = "left", 
        legend.background = element_rect(fill="grey90"), 
        legend.text = element_text(size=10, face="bold"), 
        plot.title = element_text(size = 18, face="bold"), 
        plot.subtitle = element_text(size = 14.5), 
        plot.background=element_rect(fill="grey90"), 
        panel.grid.major.y = element_blank(), 
        panel.grid.minor.y = element_blank(), 
        panel.grid.major.x=element_line(colour="darkgrey"), 
        panel.grid.minor.x=element_blank(), 
        axis.text.y = element_text(size=15)) + 
  scale_y_continuous(position="right", limits = c(0, 25.5), expand = c(0,0)) 
```


# Drivers Involved In Fatal Collisions Who Had Not Been Involved In Any Previous Accidents
```{r, echo=FALSE}
NP_data = FC_data %>% filter(type == "num_no_previous" | type == "num_previous")
NP_color = c(num_no_previous = "#269E14", num_previous= "#99C592") #darkgreen & lightgreen

ggplot(data = NP_data, aes(x = fct_rev(state), y = value, fill = fct_rev(type))) + 
  coord_flip() + labs(x= NULL, y = NULL) + geom_col(width = 0.9) + 
  ggtitle("Drivers Involved In Fatal Collisions Who Had Not \nBeen Involved In Any Previous Accidents", 
          subtitle = "As a share of the number of fatal collisions per billion miles, 2012") + 
  scale_fill_manual(labels = c("COLLISIONS PER BILLION MILES OF DRIVERS W/O PREVIOUS ACCIDENTS", "TOTAL COLLISIONS PER BILLION MILES"), 
                    values = NP_color, name = NULL) + 
  theme(legend.direction = "vertical", 
        legend.position = "top",
        legend.justification = "left", 
        legend.background = element_rect(fill="grey90"), 
        legend.text = element_text(size=10, face="bold"), 
        plot.title = element_text(size = 18, face="bold"), 
        plot.subtitle = element_text(size = 14.5), 
        plot.background=element_rect(fill="grey90"), 
        panel.grid.major.y = element_blank(), 
        panel.grid.minor.y = element_blank(), 
        panel.grid.major.x=element_line(colour="darkgrey"), 
        panel.grid.minor.x=element_blank(), 
        axis.text.y = element_text(size=15)) + 
  scale_y_continuous(position="right", limits = c(0, 25.5), expand = c(0,0)) 

```

# Drivers Involved In Fatal Collisions While Speeding
```{r, echo=FALSE}
NS_data = FC_data %>% filter(type == "num_no_speeding" | type == "num_speeding")
NS_color = c(num_speeding ="#FA1212",num_no_speeding = "#F7A1A1")

ggplot(data = NS_data, aes(x = fct_rev(state), y = value, fill = type)) + 
  coord_flip() + 
  labs(x= NULL, y = NULL) + 
  geom_col(width = 0.9) + 
  ggtitle("Drivers Involved In Fatal Collisions While Speeding", 
          subtitle = "As a share of the number of fatal collisions per billion miles, 2009") + 
  scale_fill_manual(labels = c("SPEEDING COLLISIONS PER BILLION MILES", "TOTAL COLLISIONS PER BILLION MILES"), 
                    values = NS_color, name = NULL) + 
  theme(legend.direction = "vertical", 
        legend.position = "top",
        legend.justification = "left", 
        legend.background = element_rect(fill="grey90"), 
        legend.text = element_text(size=10, face="bold"), 
        plot.title = element_text(size = 18, face="bold"), 
        plot.subtitle = element_text(size = 14.5), 
        plot.background=element_rect(fill="grey90"), 
        panel.grid.major.y = element_blank(), 
        panel.grid.minor.y = element_blank(), 
        panel.grid.major.x=element_line(colour="darkgrey"), 
        panel.grid.minor.x=element_blank(), 
        axis.text.y = element_text(size=15)) + 
  scale_y_continuous(position="right", limits = c(0, 25.5), expand = c(0,0)) 

```

# Drivers Involved In Fatal Collisions While Alcohol-Impaired
```{r, echo=FALSE}
NA_data = FC_data %>% filter(type == "num_no_alchol" | type == "num_alchol")
NA_color = c(num_alchol = "#1888F8", num_no_alchol = "#A3D5F5")

ggplot(data = NA_data, aes(x = fct_rev(state), y = value, fill = fct_rev(type))) + 
  coord_flip() + 
  labs(x= NULL, y = NULL) + 
  geom_col(width = 0.9) + 
  ggtitle("Drivers Involved In Fatal Collisions While Alcohol- \nImpaired", 
          subtitle = "As a share of the number of fatal collisions per billion miles, 2012") + 
  scale_fill_manual(labels = c("ALCOHOL-RELATED COLLISIONS PER BILLION MILES", "TOTAL COLLISIONS PER BILLION MILES"), 
                    values = NA_color, name = NULL) + 
  theme(legend.direction = "vertical", 
        legend.position = "top",
        legend.justification = "left", 
        legend.background = element_rect(fill="grey90"), 
        legend.text = element_text(size=10, face="bold"), 
        plot.title = element_text(size = 18, face="bold"),
        plot.subtitle = element_text(size = 14.5), 
        plot.background=element_rect(fill="grey90"), 
        panel.grid.major.y = element_blank(), 
        panel.grid.minor.y = element_blank(), 
        panel.grid.major.x=element_line(colour="darkgrey"), 
        panel.grid.minor.x=element_blank(), 
        axis.text.y = element_text(size=15)) + 
  scale_y_continuous(position="right", limits = c(0, 25.5), expand = c(0,0)) 
```

# Average Combined Insurance Premium
```{r, echo=FALSE}
IP_data = FC_data %>% filter(type == "premium")
IP_data$value = round(IP_data$value, digits=0)
IP_color = c(premium = "#6C1A6B")

ggplot(data = IP_data, aes(x = fct_rev(state), y = value, fill = type)) + 
  coord_flip() + 
  labs(x= NULL, y = NULL) + 
  geom_col(width = 0.9) + 
  geom_text(aes(label = value), colour= "white", size=7, hjust=1.75) +
  ggtitle("Average Combined Insurance Premium", subtitle = "2011") + 
  scale_fill_manual(values = IP_color, name = NULL) + 
  theme(legend.position="none",
        plot.title = element_text(size = 18, face="bold"), 
        plot.subtitle = element_text(size = 14.5), 
        plot.background=element_rect(fill="grey90"), 
        panel.grid.major.y = element_blank(), 
        panel.grid.minor.y = element_blank(), 
        panel.grid.major.x=element_blank(), 
        panel.grid.minor.x=element_blank(), 
        axis.text.y = element_text(size=15), 
        axis.ticks.x = element_blank(), 
        axis.text.x = element_blank()) + 
  scale_y_continuous(position="right", expand = c(0,0)) 
 
```

# Losses Incurred By Insurance Companies
```{r, echo=FALSE}
Il_data = FC_data %>% filter(type == "loss" )
Il_data$value = round(Il_data$value, digits=0)
Il_color = c(loss = "#E46A9E")

ggplot(data = Il_data, aes(x = fct_rev(state), y = value, fill = type)) + 
  coord_flip() + 
  labs(x= NULL, y = NULL) + 
  geom_col(width = 0.9) + 
  geom_text(aes(label = value), colour= "white", size=7, hjust=1.75) +
  ggtitle("Losses Incurred By Insurance Companies", subtitle = "Collisions per insured driver, 2010") + 
  scale_fill_manual(values = Il_color, name = NULL) + 
  theme(legend.position="none",
        plot.title = element_text(size = 18, face="bold"), 
        plot.subtitle = element_text(size = 14.5), 
        plot.background=element_rect(fill="grey90"), 
        panel.grid.major.y = element_blank(), 
        panel.grid.minor.y = element_blank(), 
        panel.grid.major.x=element_blank(), 
        panel.grid.minor.x=element_blank(), 
        axis.text.y = element_text(size=15), 
        axis.ticks.x = element_blank(), 
        axis.text.x = element_blank()) + 
  scale_y_continuous(position="right", expand = c(0,0)) 
```


## Technical Discussion
In this part, I will start a discussion on which steps and functions I mainly use specifically, and how they work to create my visualization plots. At the beginning, I set the display size of the images, access the library to be used, read the data and classify them, which are quite easy.  In the graphic recreation part, the most important functions are ggplot(), coord_flip() and theme(). ggplot() and coord_flip() mainly provides a mapping of data to pictures and flips them, which are the foundation of all. theme() Literally sets the theme of the image, such as background, axis, legend, title font, etc. It is because of theme() that the picture is restored as much as possible. Other functions are more auxiliary such as providing titles and setting width, and I use additionally geom_text() to display the values on the picture for the last two pictures. It is not difficult to display data in visualization project, where the recreation of details matters. In the display of histograms, because I didn't fully understand the context of each data type, I used fct_rev() incorrectly, resulting in strange output, which bothered me for hours. For the color election, I used this external webiste: https://htmlcolorcodes.com/


## Context Explaination
I recreated all the six data graphs in the original article. The first four pictures are relatively complex, while the last two pictures are relatively simple.However, their overall establishment process is similar, and the specific technology use has been introduced in the previous section. The first four pictures compare the overall number of fatal crashes per billion miles to those fatal collisions that had a particular factor present.These four factors are: Whether the driver is distracted; Whether the driver has a record before; Whether the driver exceeds the speed limit; Whether the driver drinks alcohol. The last two pictures show relatively directly the average insurance price paid by the driver every year and the amount paid by the insurance company for each collision.The data in the above six pictures show the individual data of each state. To sum up, Idaho drivers are relatively good, while New Jersey drivers are worse based on their behaviors.


## Reference
Article Title: Dear Mona, Which State Has The Worst Drivers?
Hyperlink: https://fivethirtyeight.com/features/which-state-has-the-worst-drivers/

